
\chapter{Introduction}

This report details the design and implementation of a distributed computing system to solve exact cover problems.
Donald Knuth's Dancing Links (DLX) algorithm \cite{knuth00dancing} is used to solve the exact cover problem.
Distributed computing with this algorithm is accomplished by making use of the recursive nature of DLX to split the problem into smaller pieces.
The Generic Distributed Exact Cover Solver (DECS) then takes advantage of a distributed computing middleware called BOINC \cite{boinc} to handle the work distribution and result collection process.
The report explains in detail how the DLX algorithm works and some concrete types of problems it can be applied to.
It also explains how DLX is used together with BOINC to construct a complete distributed computing system.



\section{Related work}

There are numerous implementations of the Dancing Links algorithm available with and without source code.
In addition to Knuth's own implementation in CWEB \cite{cweb} a quick search found source code for Java, Python, C, C++, Ruby, Lisp, MATLAB and Mathematica.
Some of the implementations were generic while others aimed for a specific application (mostly Sudoku).
Common for all these implementations is that none of them were made for parallel processing.

Alfred Wassermann developed a parallel version of Knuth's algorithm in \cite{wassermann99covering} by using PVM \cite{pvm} to solve a problem presented in Knuth's original paper.
However, he only published the solutions to the problem and not the actual implementation.
The only available open source parallel version is written by Owen O'Malley for the Apache Hadoop project \cite{hadoop} in May 2007.
O'Malley's implementation uses the MapReduce framework \cite{map-reduce} provided by Hadoop to do the computations in parallel.
The details of this implementation and how it divides the problem into smaller pieces has not been documented.
%The documentation for this implementation is lacking and it is pretty ...
% TODO Maybe some reasons here why O'Malley's implementation wasn't good enough, or had the required amound of academical rigor.



\section{Report organization}

This report is organized into several chapters.
Chapter \ref{dancing_links} describes the Dancing Links algorithm.
Chapter \ref{distributed_computing} explains the distributed computing concept.
Chapter \ref{implementation} discusses different aspects of the implementation.
Chapter \ref{testing} describes the test results with the system and Chapter \ref{conclusion} concludes this report.




